---
title: "BC1_PI.BF.eR"
author: "FLC"
date: '2022-05-11'
output: html_document
---
```{r}
library(ggplot2)
library(gridExtra)
library(reshape2)
library(mixOmics)
library(RColorBrewer)
library(VennDiagram)

library(ggrepel)
library(stringr)

library(edgeR)  # What will make the magic happen

library(readODS)
library(readxl)

library(tidyr)
library(dplyr)
```


# **Metadata** loading
```{r load_metadata}
# Loading of 'Animals.csv'
Animals <- read.delim("/home/fllobet/Documents/RNA-Seq_TFM/Animals.csv", sep =";", row.names = 3)

# Addition of the 'X' to the row names of Aniamls to  match col names of the
# counts matrix
rownames(Animals) <- paste("X", rownames(Animals), sep = "")

# Selection of 'BC1_PI'
Animals.BC1_PI <- Animals[which(Animals[,2] == "BC1_PI"),]

# Selection of 'liver'
Animals.BC1_PI.BF <- Animals.BC1_PI[which(Animals.BC1_PI[,3] == "Backfat"),]
```

# **Counts** loading

The lodad counts are the *integer counts* extracted from the corresponding _DESeq2 script_ where they were genearted by means of the `tximport` package.
```{r load_counts}
cnt.BC1_PI.BF <- read.delim("/home/fllobet/Documents/RNA-Seq_TFM/BC1_PI.BF/COUNTS.BC1_PI.backfat.csv", sep = ";", row.names = 1)

# Conversion into a matrix
cnt.BC1_PI.BF <- as.matrix(cnt.BC1_PI.BF)
```

# Creation of a `DGEList` **object**
Like in `desEq2` where a `dds` object had to be created here a `DGEList` object must be created.
```{r DGEList_object}
compare <- relevel(as.factor(Animals.BC1_PI.BF$Group), ref = "L")

################ CHECKING COLS count matrix == ROWS metadata ###################
# The while loop serves to see the OK prompt when the orders has been modified. 
while (TRUE) {
  if (levels(as.factor(rownames(Animals.BC1_PI.BF) == 
                       colnames(cnt.BC1_PI.BF))) == TRUE) {
    # If all the rows == all the columns (only TRUE): 
    print("The columns of the counts matrix are the same as 
          the rows of the metadata")
    break # Exit infinite loop. 
} else {
  # Change the order of the columns of the count matrix to match the metadata
  # which is the order of the "COMPARE" vector. 
  cnt.BC1_PI.BF <- cnt.BC1_PI.BF[,rownames(Animals.BC1_PI.BF)]
}
}

dge.BC1_PI.BF <- DGEList(cnt.BC1_PI.BF, group = compare)
dge.BC1_PI.BF
```



# Filtering
The native `edgeR` function `filterByExpr()`. 
This function filters features that have **CPM above _k_ in _n_ samples** where:
- _n_ is the smallest group size (in our case both groups have same size)
- _k_ is the CPM equal to 10 / _L_ where _L_ is the **minimum library size**

In human words it will remove any feature that has more than 10 counts _(but normalizing by library depth)_ in at least > samples than the small group.
```{r filtering}
a <- dim(dge.BC1_PI.BF$counts)[1]

# Filtering
keep <- filterByExpr(dge.BC1_PI.BF)
dge.BC1_PI.BF <- dge.BC1_PI.BF[keep, , keep.lib.sizes=FALSE]

# Percentage of remove features
round(((dim(dge.BC1_PI.BF$counts)[1] - a) / a), 4)

# Re-computation of the library sizes
dge.BC1_PI.BF$samples$lib.size <- colSums(dge.BC1_PI.BF$counts)
```
# Normalization - Computation of **Normalization Factors**
The **normalization factors** take into account **library size** via the *TMM* method. These factors will be used later in the `edgeR`model, the used values within the model are analogous to **CPM**.

The Normalization Factors are **considered automatically** into the FC and p.value calculations. 
```{r normalization_factors}
# Demonstration of the variability in library size between samples
st <- expression(paste("Note the ", bold("heterogeneity"), " and hence the ", bold("importance of normalization")))

DF <- cbind(rownames(dge.BC1_PI.BF$samples), dge.BC1_PI.BF$samples)
colnames(DF) <- c("Sample", colnames(DF[2:ncol(DF)]))

# Saving the object 'DF' to be used for later creation of lib sizes plots for 
# the thesis
saveRDS(DF, file = "/home/fllobet/Documents/RNA-Seq_TFM/Extra_figures/eR_LIB/BC1_PI.BF.rds")

ggplot(DF[order(Animals.BC1_PI.BF$Group),], aes(x= `rownames(dge.BC1_PI.BF$samples)`, y = lib.size*1e-6, fill = group)) + geom_bar(stat = "identity", col = "#4d5d53") + labs(title = "Libray size accross samples", subtitle = st, x = "Sample", y = "Library size (millions)") + scale_x_discrete(breaks = rownames(dge.BC1_PI.BF$samples[order(Animals.BC1_PI.BF$Group),]), labels = str_replace(str_replace(rownames(dge.BC1_PI.BF$samples[order(Animals.BC1_PI.BF$Group),]), "X", ""),"_BF", "")) + 
scale_fill_manual(name = "Group", values = c("#bdb76b","#2f847c")) + theme_bw()

# Computation of normalization factors (accounting for library size)
dge.BC1_PI.BF$samples <- calcNormFactors(dge.BC1_PI.BF)
dge.BC1_PI.BF$samples

# Readdition of the group data beause it gets removed in the previous step
dge.BC1_PI.BF$samples$group <- compare 

################## CHECKING COLS DGE object == ROWS metadata ###################
# The while loop serves to see the OK prompt when the orders has been modified. 
while (TRUE) {
  if (levels(as.factor(rownames(Animals.BC1_PI.BF) == 
                       colnames(dge.BC1_PI.BF))) == TRUE) {
    # If all the rows == all the columns (only TRUE): 
    print("The columns of the counts matrix are the same as 
          the rows of the metadata")
    break # Exit infinite loop. 
} else {
  # Change the order of the columns of the count matrix to match the metadata
  # which is the order of the "COMPARE" vector. 
  cnt.BC1_PI.BF <- cnt.BC1_PI.BF[,rownames(Animals.BC1_PI.BF)]
}
}

```
# Disersion estimation
Following the advice of DCP, we use **only `common` dispersion** so the significant genes will be those that with **extremely large counts in some libraries (samples) and not all the libraries (samples)**.
```{r common_dispersion}
dge.BC1_PI.BF <- estimateCommonDisp(dge.BC1_PI.BF)

# Value of the common dispersion:
dge.BC1_PI.BF$common.dispersion
sqrt(dge.BC1_PI.BF$common.dispersion)
```

# Exploratory data analysis

## Generation of CPM
```{r CPM}
CPM.BC1_PI.BF <- cpm(dge.BC1_PI.BF)
```

## Generation of **_pseudoCPM_**
```{r eda_01_pseudocounts}
pseudoCnt.BC1_PI.BF <- log2(CPM.BC1_PI.BF + 1)
head(pseudoCnt.BC1_PI.BF)
```

## Count distribution example
```{r eda_03_cnt_distro}
# Generation of a data frame containing 'Counts' and 'pseudoCounts' of sample 1 
df <- data.frame(counts = dge.BC1_PI.BF$counts[ ,1], 
                 pseudo = pseudoCnt.BC1_PI.BF[ ,1])

p <- ggplot(data=df, aes(x = counts, y = ..density..)) +
  geom_histogram(color = "#4d5d53", fill = "#ace1af") + theme_bw() +
  ggtitle(paste0("Count distribution '", 
                 Animals.BC1_PI.BF$Animal[1], "_LD'")) +
  xlab("Counts") + ylab("")

p2 <- ggplot(data=df, aes(x = pseudo, y = ..density..)) +
  geom_histogram(color = "#4d5d53", fill = "#2f847c") + theme_bw() +
  ggtitle(paste0("Pseudocount distribution '", 
                 Animals.BC1_PI.BF$Animal[1], "_LD'")) +
  xlab(expression(log[2](counts + 1)))

grid.arrange(p, p2, ncol = 2)


```
## ALl samples count distribution
```{r all_sample_counts}
# Merge of the metadata with the pseudocounts data
df_all_pseudoCnt <- cbind(Animals.BC1_PI.BF, t(pseudoCnt.BC1_PI.BF))

# Melting to have a 'ggplot2' ready dataframe containg metada + pseudocounts
df_all_pseudoCnt <- melt(df_all_pseudoCnt, id.vars = c("Animal", "Backcross", "Tissue", "Sex", "Group", "Father", "Mother", "Paper"))
df_all_pseudoCnt$Animal <- as.character(df_all_pseudoCnt$Animal)


p <- ggplot(data = df_all_pseudoCnt, aes(x = Animal, y = value, fill = Group)) + 
  geom_boxplot() + ggtitle("count distributions") + 
  scale_fill_manual(labels = c("H", "L"), values = c("#2f847c", "#bdb76b")) +
  xlab("sample") + ylab("counts") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90))
p
```

## MA-Plot: Reproducibility between samples
```{r eda_05_MAplot}
# Manual MA-Plot
df_ma <- data.frame("A" = (pseudoCnt.BC1_PI.BF[,1] + pseudoCnt.BC1_PI.BF[, 2])/2,
                    "M" = pseudoCnt.BC1_PI.BF[,1] - pseudoCnt.BC1_PI.BF[,2])
p <- ggplot(data = df_ma, aes(x = A, y = M)) + geom_point(alpha = 0.2, color = "#4d5d53") + theme_bw()
p
```
## Overdispersion
```{r eda_06_overdispersion}

df_ovd <- data.frame(mean = apply(cnt.BC1_PI.BF[, Animals.BC1_PI.BF$Group == "L"], 1, mean),
                     var = apply(cnt.BC1_PI.BF[, Animals.BC1_PI.BF$Group == "L"], 1, var))

df_ovd <- df_ovd[df_ovd$mean <= 5000, ]

p_ovd <- ggplot(df_ovd, aes(x = mean, y = var)) + geom_point(color = "#ace1af") + 
  theme_bw() + geom_abline(aes(intercept=0, slope=1)) + ggtitle("Variance versus means in counts of LOW group") + ylab("Variance")
p_ovd
```
For most genes the variance is higher than the mean, that is the common **overdispersionm** of RNA-Seq data. 

## Heatmap
```{r eda_07_euclidHeatmap}
# Creation of a matrix of the euclidean distances of the pseudo-counts data. 
df_hm <- as.matrix(dist(t(pseudoCnt.BC1_PI.BF)))
# Colors of the heatmap. 
cim_color <- colorRampPalette(rev(brewer.pal(9, "YlGnBu")))(16)

cim(df_hm, color = cim_color, symkey = FALSE)

```

## PCA
```{r eda_08_pca}

# Selection of the top 500 SD features
# SD for all the rows
rowSD <- apply(pseudoCnt.BC1_PI.BF, 1, sd)
# Ordering by SD
pseudoCnt.BC1_PI.BF.top500 <- pseudoCnt.BC1_PI.BF[order(rowSD,
                                                        decreasing = TRUE),]
# Selection of the top 500 more variable features
pseudoCnt.BC1_PI.BF.top500 <- pseudoCnt.BC1_PI.BF.top500[1:500,]
################################################################################
# Computation of the PCA
resPCA <- prcomp(pseudoCnt.BC1_PI.BF)
summary(resPCA)
resPCA$sdev
# Percentages of variation explained per PC
pov <- resPCA$sdev^2/sum(resPCA$sdev^2)
# Rounding in percentages:
pov <- (round(pov, digits = 4) * 100)

################################################################################
# Conserving the Principal domponents data (eigenvalues or eigenvectors?)
resPCA_pc12 <- as.data.frame(resPCA$rotation)
# Addition of the sample metadata
resPCA_pc12 <- cbind(resPCA_pc12, Animals.BC1_PI.BF)

rlg <- expression(paste(bold("pseudocounts"), " (log2(CPM + 1)) transformation applied"))
capt <- expression(paste("Counts of ", bold("liver"), " samples of ", bold("BC1_PI")))

PCA.plot <- ggplot(resPCA_pc12, aes(x = PC1, y = PC2, color = Group)) + 
  geom_point(size=3) + 
  geom_text_repel(aes(label = str_replace(
    str_replace(rownames(Animals.BC1_PI.BF), "_LD", ""),"X", ""))) +
  xlab(paste0("PC1: ",pov[1],"% variance")) +
  ylab(paste0("PC2: ",pov[2],"% variance")) + 
  scale_color_manual(values = c("#2f847c", "#bdb76b")) +  
  labs(title = "PCA of the top 500 more variable features",  
       subtitle = rlg, caption = capt) + theme_bw()

# Selection of the top 500 SD features
# SD for all the rows
rowSD <- apply(pseudoCnt.BC1_PI.BF, 1, sd)
# Ordering by SD
pseudoCnt.BC1_PI.BF.top500 <- pseudoCnt.BC1_PI.BF[order(rowSD,
                                                        decreasing = TRUE),]
# Selection of the top 500 more variable features
pseudoCnt.BC1_PI.BF.top500 <- pseudoCnt.BC1_PI.BF.top500[1:500,]
################################################################################
# Computation of the PCA
resPCA <- prcomp(pseudoCnt.BC1_PI.BF)
summary(resPCA)
resPCA$sdev
# Percentages of variation explained per PC
pov <- resPCA$sdev^2/sum(resPCA$sdev^2)
# Rounding in percentages:
pov <- (round(pov, digits = 4) * 100)

################################################################################
# Conserving the Principal domponents data (eigenvalues or eigenvectors?)
resPCA_pc12 <- as.data.frame(resPCA$rotation)
# Addition of the sample metadata
resPCA_pc12 <- cbind(resPCA_pc12, Animals.BC1_PI.BF)

rlg <- expression(paste(bold("pseudocounts"), " (log2(CPM + 1)) transformation applied"))
capt <- expression(paste("Counts of ", bold("liver"), " samples of ", bold("BC1_PI")))

PCA.plot <- ggplot(resPCA_pc12, aes(x = PC1, y = PC2, color = Group)) + 
  geom_point(size=3) + 
  geom_text_repel(aes(label = str_replace(
    str_replace(rownames(Animals.BC1_PI.BF), "_LD", ""),"X", ""))) +
  xlab(paste0("PC1: ",pov[1],"% variance")) +
  ylab(paste0("PC2: ",pov[2],"% variance")) + 
  scale_color_manual(values = c("#2f847c", "#bdb76b")) +  
  labs(title = "PCA of the top 500 more variable features",  
       subtitle = rlg, caption = capt) + theme_bw()

PCA.plot
# Obtaining the values from the MDS plot function
mds <- plotMDS(pseudoCnt.BC1_PI.BF.top500, plot = FALSE)

# Addition of sample names
mds.eigen <- as.data.frame(mds$eigen.vectors[,1:2])
rownames(mds.eigen) <- rownames(mds$distance.matrix.squared)

# Re-sorting of the 'mds.eigen' object
mds.eigen <- mds.eigen[rownames(Animals.BC1_PI.BF),]

# Addition of group data
mds.eigen <- cbind(mds.eigen, Animals.BC1_PI.BF[order(rownames(mds.eigen)),7])
colnames(mds.eigen) <- c(colnames(mds.eigen)[1:2], "Group")
# Manual plotting of the eigenVectors form the MDS plot output
ggplot(mds.eigen, aes(x = V1, y = V2, color = Group, 
                      label = str_replace(str_replace(rownames(mds.eigen), 
                                                      "X", ""),"_BF", ""))) + 
  geom_text() + 
  scale_color_manual(values = c(rgb(47/225,132/225,124/225), 
                                rgb(189/225,183/225,107/225))) + 
  labs(x = paste0("Leading logFC dim 1 (", round(mds$var.explained[1]*100,0), 
                  "%)"),
       y = paste0("Leading logFC dim 2 (", round(mds$var.explained[2]*100,0),
                  "%)"), 
       title = "BC1_PI ~ BF", 
       subtitle = "MDS plot (PCA) of the top 500 more variable features",
       caption = "Using root-mean-square average of the top log2FC between samples (leading log2FC).") + 
  guides(
  color = guide_legend(title = "Group", override.aes = aes(label = "•", 
                                                           size = 8))) +
  scale_y_reverse() + scale_x_reverse() + 
  theme_bw()
# plotMDS(pseudoCnt.BC1_PI.BF.top500, col = c(rgb(189/225,183/225,107/225),
#                                             rgb(47/225,132/225,124/225),
#                                             rgb(189/225,183/225,107/225),
#                                             rgb(189/225,183/225,107/225),
#                                             rgb(47/225,132/225,124/225),
#                                             rgb(47/225,132/225,124/225),
#                                             rgb(189/225,183/225,107/225),
#                                             rgb(47/225,132/225,124/225),
#                                             rgb(189/225,183/225,107/225),
#                                             rgb(189/225,183/225,107/225),
#                                             rgb(47/225,132/225,124/225),
#                                             rgb(47/225,132/225,124/225)), labels = str_replace(str_replace(colnames(pseudoCnt.BC1_PI.BF.top500), "X", ""),"_LV", ""))
# legend("bottomright", legend = c("H", "L"), col = c(rgb(47/225,132/225,124/225), rgb(189/225,183/225,107/225)), pch = 19)
# title(main = "MDS plot (PCA) of the top 500 more variable features", sub = "Using root-mean-square average of the top log2FC between samples (leading log2FC).", cex.sub = 0.8)
```
# Differential expression analysis: Pairwise exact test
The test is quite similar to the Fisher's exact test
```{r pairwise_exact_test}
# Lodaing of the file where I have the list of significant features from 
# 'Corominas et al. 2013'
coro.BC1_LD.BF <- read_excel("/home/fllobet/Documents/RNA-Seq_TFM/BC1_LD.BF/DEG_Corominas_2013_a.05_FC_1.2.xlsx")
colnames(coro.BC1_LD.BF) <- c("ENSEMBL ID", "Gene name", colnames(coro.BC1_LD.BF)[3:4])
# Exact test (core statistical test)
res_BC1_PI.BF <- exactTest(dge.BC1_PI.BF)
# Vislualitzation of the top 10 most significant features
topTags(res_BC1_PI.BF)

################################################################################
# FDR Correction: Benjamini-Honchberg.
res_BC1_PI.BF$table <- 
  cbind(res_BC1_PI.BF$table, 
        p.adjust(res_BC1_PI.BF$table$PValue, method = "BH"))

# Confirmation of the performed comparison
res_BC1_PI.BF$comparison

################################################################################
# Storing of the results into a CSV file
res_BC1_PI.BF$table <- cbind(rownames(res_BC1_PI.BF$table), res_BC1_PI.BF$table)
res_BC1_PI.BF$table <- cbind(rownames(res_BC1_PI.BF$table), res_BC1_PI.BF$table)
res_BC1_PI.BF$table[,1:2] <- str_split_fixed(rownames(res_BC1_PI.BF$table),"_", n = 2) 

colnames(res_BC1_PI.BF$table) <- c("ENSEMBL ID", "Gene name", colnames(res_BC1_PI.BF$table)[3:6])

# BUSCARV over the lists - ENSEMBL ID

IDs_in_coro <- merge(res_BC1_PI.BF$table, coro.BC1_LD.BF, by="ENSEMBL ID")$`ENSEMBL ID`

# Printing of the number of mathcing features
length(IDs_in_coro)

# Addition of this info in the output CSV of the significant genes
L <- dim(res_BC1_PI.BF$table)[2]
for (i in 1:dim(res_BC1_PI.BF$table)[1]) {
  if (res_BC1_PI.BF$table[i,1] %in% IDs_in_coro) {
    res_BC1_PI.BF$table[i,L+1] <- "in Corominas et al. 2013"
    }
}
################################################################################
# BUSCARV over the lists - GENE NAME


GNs_in_coro <- merge(res_BC1_PI.BF$table, coro.BC1_LD.BF, by="Gene name")$`Gene name`
# Printing of the number of mathcing features
length(GNs_in_coro)

L <- dim(res_BC1_PI.BF$table)[2]
for (i in 1:dim(res_BC1_PI.BF$table)[1]) {
  if (res_BC1_PI.BF$table[i, 2] %in% GNs_in_coro) {
    res_BC1_PI.BF$table[i, L + 1] <- "in Corominas et al. 2013"
  }
}

################################################################################
# Merging of the 2 columns recently created 'Paper ID' and 'Paper Gene'
colnames(res_BC1_PI.BF$table) <- c(colnames(res_BC1_PI.BF$table)[1:5], "PADJValue", "Paper ID", "Paper Gene")

# Merging the 2 columns of 'Paper ID' and 'Paper Name' to paint the plot
res_BC1_PI.BF$table  <- res_BC1_PI.BF$table %>% mutate(`Paper ID` = coalesce(`Paper ID`,`Paper Gene`)) %>%
         select(colnames(res_BC1_PI.BF$table)[1:7])



################################################################################
# Storing the results into a CSV file 
colnames(res_BC1_PI.BF$table) <- c(colnames(res_BC1_PI.BF$table)[1:5], "PADJValue", "Paper")
write.table(res_BC1_PI.BF$table, "/home/fllobet/Documents/RNA-Seq_TFM/BC1_PI.BF/edgeR/BC1_PI.BF_eResults.csv", sep = ";", row.names = FALSE)
```

```{r signficant_features}
# Return of significant features with a given alpha threshold:
signif_BC1_PI.BF <- decideTestsDGE(res_BC1_PI.BF, adjust.method = "BH", p.value = 0.05)

# Selection of the significant features
sel_signif_BC1_PI.BF <- which(signif_BC1_PI.BF[,1] != 0)

# Datafame of the Results + P.adjusted values of only the significant features
signif_BC1_PI.BF <- cbind(res_BC1_PI.BF$table[sel_signif_BC1_PI.BF,],
                          signif_BC1_PI.BF[sel_signif_BC1_PI.BF, 1])

################################################################################
# Printing of:
# 'n' of signficant features (rows)
dim(signif_BC1_PI.BF)[1]

# 'n' of logFC > 0 
dim(signif_BC1_PI.BF[which(signif_BC1_PI.BF$logFC > 0 & signif_BC1_PI.BF$PADJValue < 0.05),])[1]
# 'n' of logFC < 0
dim(signif_BC1_PI.BF[which(signif_BC1_PI.BF$logFC < 0 & signif_BC1_PI.BF$PADJValue < 0.05),])[1]

# Storing the signficiant features into a CSV file
write.table(signif_BC1_PI.BF, "/home/fllobet/Documents/RNA-Seq_TFM/BC1_PI.BF/edgeR/BC1_PI.BF_eResults_Signif.05.csv", sep = ";", row.names = FALSE)
```

# ABS Fold-Change > 1.2
```{r fc_1.2}
fc12 <- log2(1.2)

# Filtering 
signif_BC1_PI.BF_FC_1.2 <- signif_BC1_PI.BF[which(signif_BC1_PI.BF$logFC > fc12 | signif_BC1_PI.BF$logFC < -fc12),]
dim(signif_BC1_PI.BF_FC_1.2)[1]

# 'n' of logFC > 0 
dim(signif_BC1_PI.BF_FC_1.2[which(signif_BC1_PI.BF_FC_1.2$logFC > 0 & signif_BC1_PI.BF_FC_1.2$PADJValue < 0.05),])[1]
# 'n' of logFC < 0
dim(signif_BC1_PI.BF_FC_1.2[which(signif_BC1_PI.BF_FC_1.2$logFC < 0 & signif_BC1_PI.BF_FC_1.2$PADJValue < 0.05),])[1]
```
# ABS Fold-Change > 1.5
```{r fc_1.5}
fc15<- log2(1.5)

# Filtering 
signif_BC1_PI.BF_FC_1.5 <- signif_BC1_PI.BF[which(signif_BC1_PI.BF$logFC > fc15 | signif_BC1_PI.BF$logFC < -fc15),]
dim(signif_BC1_PI.BF_FC_1.5)[1]

# 'n' of logFC > 0 
dim(signif_BC1_PI.BF_FC_1.5[which(signif_BC1_PI.BF_FC_1.5$logFC > 0 & signif_BC1_PI.BF_FC_1.5$PADJValue < 0.05),])[1]
# 'n' of logFC < 0
dim(signif_BC1_PI.BF_FC_1.5[which(signif_BC1_PI.BF_FC_1.5$logFC < 0 & signif_BC1_PI.BF_FC_1.5$PADJValue < 0.05),])[1]
```
# PValues

```{r}
levels(as.factor(res_BC1_PI.BF$table$Paper))
ggplot(res_BC1_PI.BF$table[which(res_BC1_PI.BF$table$PADJValue < 1),], 
       aes(x = PADJValue, fill = as.factor(Paper))) + 
  geom_histogram(color = "#4d5d53") + 
  geom_vline(xintercept = 0.05, linetype='dashed', col = '#556b2f', size = 1) + 
  geom_vline(xintercept = 0.1, linetype='dashed', col = '#2f847c', size = 1) +
  scale_fill_manual(name = "", values = c("#ace1af"), na.value = "#bdb76b", 
                    labels = c("in Corominas et al. 2013", 
                               "NOT in Corominas et al. 2013")) + 
  labs(x = "Adjusted PValues distribution", y = "Frequency", title = "PValues distribution of BC1_PI ~ BACKFAT", subtitle = "Location of the signficiant features of Corominas et al. 2013 alognside the adjuste PValue distribution", caption = "The dashed line repsent 0.05 and 0.1 respectively") + theme_bw() + 
  theme(legend.position = "bottom")
```


# Saving the CPM table
```{r save_CPM}
# Ordering L -> H the columns
CPM.BC1_PI.BF <- CPM.BC1_PI.BF[,rownames(Animals.BC1_PI.BF[order(Animals.BC1_PI.BF$Group, decreasing = TRUE),])]

# Division of the 'Features' column into 2 new columns
CPM.BC1_PI.BF <- cbind(rownames(CPM.BC1_PI.BF),
                       rownames(CPM.BC1_PI.BF),
                       CPM.BC1_PI.BF)

CPM.BC1_PI.BF[,1:2] <- str_split_fixed(rownames(CPM.BC1_PI.BF),"_", n = 2) 

colnames(CPM.BC1_PI.BF) <- c("ENSEMBL ID", "Gene name", str_replace(colnames(CPM.BC1_PI.BF)[3:dim(CPM.BC1_PI.BF)[2]],"X",""))

write.table(CPM.BC1_PI.BF, "/home/fllobet/Documents/RNA-Seq_TFM/BC1_PI.BF/edgeR/CPM.BC1_PI.BF.csv", sep = ";", row.names = FALSE) 

```