---
title: "BC1_LD.LV.IsF.EB.Rmd"
author: "FLC"
date: '2022-05-15'
output: html_document
---
```{r}
library(stringr)
#BiocManager::install("EBSeq")
library(EBSeq)
```


# Lodaing *expected counts* from RSEM
```{r rsem_Counts_LOADING}
RSEM_exCnt.BC1_LD.iF.LV <- read.delim("/home/fllobet/Documents/RNA-Seq_TFM/BC1_LD.LV/RSEM_counts_BC1.LD.LV.matrix", sep = "\t", row.names = 1)
# Modification of the colnames:
colnames(RSEM_exCnt.BC1_LD.iF.LV) <- str_replace(str_replace(colnames(RSEM_exCnt.BC1_LD.iF.LV),".isoforms.results",""),"^[.-_]+\\.","X")
View(RSEM_exCnt.BC1_LD.iF.LV)
```

# Metadata loading
```{r metadata_LOADING}
### METADATA of ALL SAMPLES
Animals <- read.delim("/home/fllobet/Documents/RNA-Seq_TFM/Animals.csv", sep = ";", stringsAsFactors = FALSE)

## Cleaning of the metadata
# Relocation of the Animal number as row names 
row.names(Animals) <- paste0("X",Animals[,3])
# The 'X' is necessary because R cannot use column names staring wiht a number.

### Subset of the METADATA from the BC1_LD liver samples:

# BC1_LD
Animals.BC1_LD <- Animals[which(Animals[,2] == "BC1_LD"),]
# Liver
Animals.BC1_LD.iF.LV <- Animals.BC1_LD[which(Animals.BC1_LD[,4] == "Liver"),]

# Ensure that both files follow the same order
Animals.BC1_LD.iF.LV <- Animals.BC1_LD.iF.LV[order(Animals.BC1_LD.iF.LV[,1]),]
RSEM_exCnt.BC1_LD.iF.LV <- RSEM_exCnt.BC1_LD.iF.LV[,rownames(Animals.BC1_LD.iF.LV)]

################################################################################
# Removal of unutilized objects
rm(Animals)
```
The group vector will be `r Animals.BC1_LD.iF.LV$Group`


# Selection of only the *female* animals
```{r coro_animalse}
# Seletion of the 6 animals from the metadata
Animals.BC1_LD.iF.LV <- Animals.BC1_LD.iF.LV[which(Animals.BC1_LD.iF.LV[,5] == "F"),]

# Seletion of the 6 animals from the counts matrix
RSEM_exCnt.BC1_LD.iF.LV <- RSEM_exCnt.BC1_LD.iF.LV[, rownames(Animals.BC1_LD.iF.LV)]
```

# Isoform and gene names
```{r isoform_names}
iso_names <- rownames(RSEM_exCnt.BC1_LD.iF.LV)
```

```{r gene_names}
# The first 'isofroms.results' file of the data will be used to retrieve the gene names
file <- paste0("/home/fllobet/Documents/RNA-Seq_TFM/COUNTS/Isoforms/PK02/",str_replace(colnames(RSEM_exCnt.BC1_LD.iF.LV)[1],"X",""),".isoforms.results")
gene_names <- read.delim(file)$gene_id
```

# Normalization: Library Size Factor
```{r library_size}
libSzF.BC1_LD.iF.LV <- MedianNorm(RSEM_exCnt.BC1_LD.iF.LV)
```

# N_{g} vector
EBSeq fits diferent prior parameters for diferent uncertainty groups (defned as Ig groups). The default setting to defne the uncertainty groups consists of using the number of isoforms the host gene contains (Ng ) for each isoform.
```{r Ng_vector}
Ng_BC1_LD.iF.LV <- GetNg(IsoformName = iso_names, GeneName = gene_names)
# The ACTUAL Ng vector that will be feed to the EBSeq model fitting
NgV_BC1.Pi.iF.LV <- Ng_BC1_LD.iF.LV$IsoformNgTrun

```

# Differential expression analysis
1 Fitting the model 
```{r EBTestT}
# Condition vector refactoring
groups <- relevel(as.factor(Animals.BC1_LD.iF.LV$Group), ref = "L")

# Usually the the 3 parameters (alfa, beta & prior) do not stabilize 
# (changing les e-3 = 0.01) until iteration 12-13, hence the deafault number
# of iterations is 15:

# Model fiting
EB.BC1_LD.iF.LV <- EBTest(Data=as.matrix(RSEM_exCnt.BC1_LD.iF.LV), 
                          NgVector=NgV_BC1.Pi.iF.LV, 
                          Conditions=groups,
                          sizeFactors=as.vector(libSzF.BC1_LD.iF.LV), 
                          maxround=15)

EB.BC1_LD.iF.LV$Alpha
EB.BC1_LD.iF.LV$Beta
EB.BC1_LD.iF.LV$P
```

2 Results (with a FDR cutoff of .05)
```{r EBResults}
# Applying the robust method
robust.EBRes.BC1_LD.iF.LV <- GetDEResults(EB.BC1_LD.iF.LV, FDR = 0.05, Method = "robust")

# Aoplying the classic method
classic.EBRes.BC1_LD.iF.LV <- GetDEResults(EB.BC1_LD.iF.LV, FDR = 0.05, Method = "classic")

# Nº of DE Isoforoms 
# Robust method
length(robust.EBRes.BC1_LD.iF.LV$DEfound)
# Classic method
length(classic.EBRes.BC1_LD.iF.LV$DEfound)
```
# Calculation of Fold-Changes
```{r FC}
FC.BC1_LD.iF.LV <- PostFC(EB.BC1_LD.iF.LV)
FC.BC1_LD.iF.LV <- FC.BC1_LD.iF.LV$RealFC
# TO GET NUMBER OF FILTERED ISOFOROMS
length(FC.BC1_LD.iF.LV)
length(classic.EBRes.BC1_LD.iF.LV$PPMat)

# Generation of a results matrix of the classic method:
classicalFC.PP_matrix <- cbind(names(FC.BC1_LD.iF.LV), names(FC.BC1_LD.iF.LV), FC.BC1_LD.iF.LV, -log2(FC.BC1_LD.iF.LV), classic.EBRes.BC1_LD.iF.LV$PPMat[which(rownames(classic.EBRes.BC1_LD.iF.LV$PPMat) %in% names(FC.BC1_LD.iF.LV)),])
classicalFC.PP_matrix[,1:2] <- str_split_fixed(rownames(classicalFC.PP_matrix),"_", n = 2) 
colnames(classicalFC.PP_matrix) <- c("ENSEMBL ID", "Isoform name", "FC", "log2FC", "PP EE", "PP DE")

# Results CSV of all the features
write.table(classicalFC.PP_matrix, "/home/fllobet/Documents/RNA-Seq_TFM/BC1_LD.LV/ISOFORMS/EBSeq/BC1_LD.iF.LV_EBcl.Results.csv", sep = ";", row.names = FALSE)
# Results CSV of only the significant features
write.table(classicalFC.PP_matrix[which(rownames(classicalFC.PP_matrix) %in% classic.EBRes.BC1_LD.iF.LV$DEfound),], "/home/fllobet/Documents/RNA-Seq_TFM/BC1_LD.LV/ISOFORMS/EBSeq/BC1_LD.iF.LV_EBcl_SIGNIF.Results.csv", sep = ";", row.names = FALSE)


# Generation of a results matrix of the robust method:
robustFC.PP_matrix <- cbind(names(FC.BC1_LD.iF.LV), names(FC.BC1_LD.iF.LV), FC.BC1_LD.iF.LV, -log2(FC.BC1_LD.iF.LV), robust.EBRes.BC1_LD.iF.LV$PPMat[which(rownames(robust.EBRes.BC1_LD.iF.LV$PPMat) %in% names(FC.BC1_LD.iF.LV)),])
robustFC.PP_matrix[,1:2] <- str_split_fixed(rownames(robustFC.PP_matrix),"_", n = 2) 
colnames(robustFC.PP_matrix) <- c("ENSEMBL ID", "Isoform name", "FC", "log2FC", "PP EE", "PP DE")
# Results CSV of all the features
write.table(robustFC.PP_matrix, "/home/fllobet/Documents/RNA-Seq_TFM/BC1_LD.LV/ISOFORMS/EBSeq/BC1_LD.iF.LV_EBrb.Results.csv", sep = ";", row.names = FALSE)

# Results CSV of only the significant features
write.table(robustFC.PP_matrix[which(rownames(robustFC.PP_matrix) %in% robust.EBRes.BC1_LD.iF.LV$DEfound),], "/home/fllobet/Documents/RNA-Seq_TFM/BC1_LD.LV/ISOFORMS/EBSeq/BC1_LD.iF.LV_EBrb_SIGNIF.Results.csv", sep = ";", row.names = FALSE)

# Results CSV of only the significant features
```
# Calculation of percentage of filtered features
```{r filtered_features}
(dim(RSEM_exCnt.BC1_LD.iF.LV)[1] - dim(EB.BC1_LD.iF.LV$PPMat)[1]) / dim(RSEM_exCnt.BC1_LD.iF.LV)[1]

47793 / 63041
```

# Up & Down regulation + Filtering by FoldChange

## ABS FC > 0
```{r abs_0}
# Only from the significant features
robustFC_signif <- robustFC.PP_matrix[which(rownames(robustFC.PP_matrix) %in% robust.EBRes.BC1_LD.iF.LV$DEfound),]

# UpRegulated: FC > 0
nrow(robustFC_signif[which(as.numeric(robustFC_signif[,4]) > 0),])

# DownRegulated: FC < 0
nrow(robustFC_signif[which(as.numeric(robustFC_signif[,4]) < 0),])
```


## ABS FC > 1.2
```{r abs_1.2}
# log2 of 1.2
lg1.2 <- log2(1.2)

# Creation of a new object

# With FC > and < 1.2
robustFC.PP_matrix.12 <- 
  robustFC_signif[which(as.numeric(robustFC_signif[,4]) > lg1.2 |  
                          as.numeric(robustFC_signif[,4]) < -lg1.2),]



# Number of DE isoforms
nrow(robustFC.PP_matrix.12)

# UpRegulated: FC > 0
nrow(robustFC.PP_matrix.12[which(as.numeric(robustFC.PP_matrix.12[,4]) > 0),])

# DownRegulated: FC < 0
nrow(robustFC.PP_matrix.12[which(as.numeric(robustFC.PP_matrix.12[,4]) < 0),])
```

## ABS FC > 1.5
```{r abs_1.5}
# log2 of 1.5
lg1.5 <- log2(1.5)


# Creation of a new object

# With FC > and < 1.5
robustFC.PP_matrix.15 <- 
  robustFC_signif[which(as.numeric(robustFC_signif[,4]) > lg1.5 |  
                          as.numeric(robustFC_signif[,4]) < -lg1.5),]



# Number of DE isoforms
nrow(robustFC.PP_matrix.15)

# UpRegulated: FC > 0
nrow(robustFC.PP_matrix.15[which(as.numeric(robustFC.PP_matrix.15[,4]) > 0),])

# DownRegulated: FC < 0
nrow(robustFC.PP_matrix.15[which(as.numeric(robustFC.PP_matrix.15[,4]) < 0),])
```
# Comparison with DESeq2 results
```{r comp_DESeq2}
BC1_LD.iF.LV_Dq2 <- read.csv("/home/fllobet/Documents/RNA-Seq_TFM/BC1_LD.LV/ISOFORMS/BC1_LD.iF.LV.SF.Results_a_0.05.csv", sep = ";")

# Selection of just the significant faetures
BC1_LD.iF.LV_Dq2 <- BC1_LD.iF.LV_Dq2[which(BC1_LD.iF.LV_Dq2[,8] < 0.05),]

# Row names of both objects become the ENSMEBL ID
rownames(BC1_LD.iF.LV_Dq2) <- BC1_LD.iF.LV_Dq2$ENSEMLB.ID
rownames(robustFC_signif) <- robustFC_signif[,1]
# 'BUSCARV' between the 2 lists
Common <- merge(robustFC_signif, BC1_LD.iF.LV_Dq2, by = 0)

# Number of common features between the 2 lists
nrow(Common)


# Addition of a new column into the output CSV of the significant features
L <- dim(robustFC_signif)[2]
robustFC_signif <- as.data.frame(robustFC_signif)
for (i in 1:nrow(robustFC_signif)) {
  robustFC_signif[i,L+1] <- ifelse(rownames(robustFC_signif[i,]) %in% Common[,1], "in DESeq2", NA)

  }

# Update of the column names
colnames(robustFC_signif) <- c(colnames(robustFC_signif)[-ncol(robustFC_signif)], "in DESeq2?")




write.table(robustFC_signif, "BC1_LD.iF.LV_EBrb_SIGNIF.Results.csv", row.names = FALSE, sep = ";")
```

# Saving the counts
Like I used to do at gene-level in DESeq2 and edgeR now I will 
```{r}
# Ordering the columns of the RSEM-estimated-counts matrix in L ... H fashion
signifRSEMCnts <- as.data.frame(RSEM_exCnt.BC1_LD.iF.LV[,
                                          colnames(RSEM_exCnt.BC1_LD.iF.LV[
                                            order(Animals.BC1_LD.iF.LV$Group, 
                                                  decreasing = TRUE)])])
# Division of the rownames into ID + transcript
signifRSEMCnts <- cbind(rownames(signifRSEMCnts), rownames(signifRSEMCnts), signifRSEMCnts)
signifRSEMCnts[,1:2] <- str_split_fixed(rownames(signifRSEMCnts),"_",n = 2)
rownames(signifRSEMCnts) <- signifRSEMCnts[,1]

# Selection of only the significant features (PP DE > 0.95)
rw <- rownames(robustFC.PP_matrix[which(as.numeric(robustFC.PP_matrix[,6])>0.95),])
signifRSEMCnts <- signifRSEMCnts[rw,]


# Saving the counts matrix into a CSV file 
colnames(signifRSEMCnts) <- c("ENSEMBL ID", "Transcript", str_replace(colnames(signifRSEMCnts[3:ncol(signifRSEMCnts)]),"X",""))
write.table(signifRSEMCnts, "/home/fllobet/Documents/RNA-Seq_TFM/BC1_LD.LV/ISOFORMS/EBSeq/BC1_LD.iF.LV_EBrb_COUNTS.csv", sep = ";", 
            row.names = F)
```


